COVID-19 and Online Learning Tools

### 2021-12-27 Paper_online_Learning.tex
working on intro section

COVID-19 is still one of the most important priorities of governments and media of many countries all around the world. Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}. As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}. In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.

@@

Many sectors have been affected ever since the outbreak of COVID-19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 \autocite{unesco2020covid}. With the start of the pandemic a near total closure of schools were observed all around the world. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).

@@

Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}. Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.

The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as teachers, students and parents. Teachers were not fully aware of their obligations and how to connect with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID-19 pandemic. Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}. This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home. Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures. Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown. This situation left internet as the only medium to support all these educational processes.

According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and need of urgent responses.

>>>>

COVID-19 is still one of the most important priorities of governments and media in many countries all around the world. Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on January 30, 2020, and then a pandemic on March 11, 2020 \autocite{world2020timeline}. As of September 29, 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}. In response to the pandemic, health authorities worldwide have taken many steps, including vaccination development and deployment to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, and reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.

@@

Many sectors have been affected ever since the outbreak of COVID-19, worldwide. Among these many sectors, education is one of the most affected sectors with a near-total closure of schools, colleges, and universities all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} show the evolution of global school closures and reopenings since mid February 2020 \autocite{unesco2020covid}. With the start of the pandemic, a near total closure of schools was observed all around the world. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are held exclusively in person (Figures \ref{fig:covidImpactContinent}).

@@

Even though distance education has a long history that goes back almost two centuries with significant modifications,
alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}. Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.

The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders, such as teachers, students, and parents. Teachers were not fully aware of their obligations and how to connect with students to support their learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, and lack of awareness of students' ongoing problems were some of the major issues they encountered during the COVID-19 pandemic. Students on the other hand who tend to have fewer educational opportunities beyond schools and universities are severely affected by this sudden and unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives, such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled devices or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}. This also put an extra burden on parents, especially those with limited education and resources, as they were expected to facilitate the required learning environment at home. Working parents also find it difficult to work during a school closure due to childcare obligations that result from unexpected school closures. Furthermore, every single step, including planning, developing new tools and techniques, conducting awareness programmes about distance education, and shifting towards distance education, happened online, due to the unexpected massive global shutdown. This situation left the internet as the only medium to support all these educational processes.

According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and the need for urgent responses.

### 2021-12-27 main.Rmd
working on intro section

COVID-19 is still one of the most important priorities of governments and  media of many countries all around the world.  Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline]. As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].  In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment  to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as  travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus [@chang2020modelling].

@@

Many sectors have been affected ever since the outbreak of COVID-19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education]. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 [@unesco2020covid]. With the start of the pandemic a near total closure of schools were observed all around the world. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).

@@

Eventhough distance education has a long history that goes back to almost two centuries with significant  modifications,
alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously [@richmond2020critical].  Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher]. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.

The speed of the physical classroom closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as  teachers, students and parents.   Teachers were not fully aware of their obligations and how to connect with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID-19 pandemic. Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher]. This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home. Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures. Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown. This situation left internet as the only medium to support all these educational processes.

According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and need of urgent responses.

>>>>

COVID-19 is still one of the most important priorities of governments and media in many countries all around the world. Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on January 30, 2020, and then a pandemic on March 11, 2020 [@world2020timeline]. As of September 29, 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].  In response to the pandemic, health authorities worldwide have taken many steps, including vaccination development and deployment  to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as  travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, and reduction of mass gatherings to reduce the spread of the virus [@chang2020modelling].

@@

Many sectors have been affected ever since the outbreak of COVID-19, worldwide. Among these many sectors, education is one of the most affected sectors with a near-total closure of schools, colleges, and universities all around the world [@daniel2020education]. Figures \ref{fig:covidImpactWorld} show the evolution of global school closures and reopenings since mid February 2020 [@unesco2020covid]. With the start of the pandemic, a near total closure of schools was observed all around the world. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are held exclusively in person (Figures \ref{fig:covidImpactContinent}).

@@

Even though distance education has a long history that goes back  almost two centuries with significant  modifications,
alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].  Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher]. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.

The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders, such as  teachers, students, and parents.   Teachers were not fully aware of their obligations and how to connect with students to support their learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, and lack of awareness of students' ongoing problems  were some of the major issues they encountered during the COVID-19 pandemic. Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities are severely affected by this sudden and unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives, such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled devices or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher]. This also put an extra burden on parents, especially those with limited education and resources,  as they were  expected to facilitate the required learning environment at home. Working parents also find it difficult to work during a school closure due to childcare obligations that result from unexpected school closures. Furthermore, every single step, including planning, developing new tools and techniques, conducting awareness programmes about distance education, and shifting towards distance education, happened online, due to the unexpected massive global shutdown. This situation left the internet as the only medium to support all these educational processes.

According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and the need for urgent responses.
